---
title: "Analysis of GC Content in Bacterial Genomes"
author: "CJA"
date: "2025-12-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(rstatix)
```
1. Introduction
This report presents an analysis of the GC content (%) in bacterial genomes across different genera.
The data were retrieved from the NCBI database and processed automatically.

2. Data Loading
```{r}
df <- read.csv("D:/Programation/Nouveau dossier/gc_genomes.csv")
df$genus <- as.factor(df$genus)
```

3. Descriptive Statistics
```{r}
desc <- df %>%
group_by(genus) %>%
summarise(
mean_gc = mean(gc_content_percent),
sd_gc = sd(gc_content_percent),
n = n()
)
desc
```
4. Normality Test
```{r}
norm_test <- df %>%
group_by(genus) %>%
summarise(shapiro_p = shapiro.test(gc_content_percent)$p.value)
norm_test
```
5. Global Statistical Test
```{r}
if(all(norm_test$shapiro_p > 0.05)){
test <- aov(gc_content_percent ~ genus, data=df)
summary(test)
posthoc <- TukeyHSD(test)
posthoc
} else {
test <- kruskal_test(gc_content_percent ~ genus, data=df)
test
posthoc <- df %>%
dunn_test(gc_content_percent ~ genus, p.adjust.method = "bonferroni")
posthoc
}
```
6. Visualization
```{r}
p <- ggplot(df, aes(x=genus, y=gc_content_percent, color=genus)) +
geom_boxplot(outlier.shape = NA) +
geom_jitter(width=0.2, alpha=0.6) +
theme_minimal(base_size = 14) +
labs(
title = "GC Content (%) by Bacterial Genus",
x = "Genus",
y = "%GC"
) +
theme(legend.position = "none")
p
```
7. Interpretation
The results highlight differences in GC content between bacterial genera.
Statistical tests and visualizations provide a clear quantitative view of these differences.
These insights can be used for comparative genomics studies or publications.

